Equalizer Programming Guide

### 2007-11-23 Developer/ProgrammingGuide/paper.tex
Add GLEW-related documentation, minor website tweaks
[ ] May break build
[ ] Breaks existing applications
[ ] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[x] Documentation

\date{
  \vspace{2cm}
  \begin{figure}[ht]
    \hspace{-4.5cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{figure}
  \vspace{1cm}
%  \textbf{RELEASE CANDIDATE}\\[\medskipamount]
  Version 1.1.2 for Equalizer 0.4+svn\\[\medskipamount]
  \today
}

@@

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.0.1   & Nov 06, 2007 & Fixed Typo\\
    1.1     & Nov 12, 2007 & Adapted distributed objects\\
    1.1.1   & Nov 16, 2007 & Minor improvements based on lulu.com print\\
    1.1.2   & Nov 19, 2007 & Minor improvements based on feedback\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/eventInit.pdf}
  {\caption{\small\label{fEventInit}Event Handling Initialization}}
\end{wrapfigure}
During window and pipe initialization the event handling is set up. For
both entities, \textsf{initEvent\-Handler} is called to register the
pipe or window with an event handler. This method may be overwritten to
use a custom event handler, or to not install an event handler at all
and so disable event handling. Likewise, \textsf{exitEventHandler} is
called to de-initialize event handling.

An event handler consists of two parts: the generic base class providing
the interface and generic functions, and the window-system-specific
part providing the actual implementation.

@@

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\small\label{fDirectSend}Parallel Direct Send Compositing}}
\end{wrapfigure}
The main idea for direct send is to parallelize the costly recomposition
for database (sort-last) decomposition. With each additional source
channel, the amount of pixel data to be composited grows linearly. When
using the simple approach of compositing all frames on the destination
channel, this channel quickly becomes the bottleneck in the
system. Direct send distributes this workload evenly across all source
channels, and thereby keeps the compositing work per channel constant.

@@

The following operations have to be possible in order to perform this
algorithm:
\begin{itemize}
\item Selection of frame buffer attachments: color and/or depth
\item Restricting the read back area to a part of the rendered area
\item Positioning the pixel data correctly on the receiving channels
\end{itemize}

@@

The frame synchronization through the frame data allows the input frame
to wait for the pixel data to become ready, which is signalled by the
output frame after readback.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/assembly.pdf}
  {\caption{\small\label{fAssembly}Hierarchy of assembly classes}}
\end{wrapfigure}
Furthermore, the frame data transports the inherited range of the output
frame's compound. The range can be used to compute the assembly order of
multiple input frames, e.g., for sorted-blend compositing in volume
rendering applications.

@@

Finding the correct assembly order for perspective frustra is more
complex. The perspective distortion invalidates a simple orientation
criteria like the one used for orthographic frustra. For the view and
frustum setup shown in \fig{fBlendPerspective}\footnote{Volume Data Set
  courtesy of: AVS, USA} the correct compositing order is 4-3-1-2 or
1-4-3-2.

\begin{wrapfigure}{r}{.382\textwidth}
  \vspace{-4ex}\includegraphics[width=.382\textwidth]{images/volResult.png}
  {\caption{\small\label{fVolResult}Result of \fig{fBlendPerspective}}}
\end{wrapfigure}
In order to compute the assembly order, \textsf{eVolve} uses the angle
between the origin$\rightarrow$slice vector and the near plane, as shown
in \fig{fBlendPerspective}. When the angle becomes greater than
90\textdegree, the compositing order of the remaining frames has to be
changed. The result image of this composition naturally looks the same
as the volume rendering would when rendered on a single
channel. \fig{fVolResult} shows the result of the composition from
\fig{fBlendPerspective}.

>>>>

\date{
  \vspace{2cm}
  \begin{figure}[ht]
    \hspace{-4.5cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{figure}
  \vspace{1cm}
%  \textbf{RELEASE CANDIDATE}\\[\medskipamount]
  Version 1.1.3 for Equalizer 0.4+svn\\[\medskipamount]
  \today
}

@@

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.0.1   & Nov 06, 2007 & Fixed Typo\\
    1.1     & Nov 12, 2007 & Adapted distributed objects\\
    1.1.1   & Nov 16, 2007 & Minor improvements based on lulu.com print\\
    1.1.2   & Nov 19, 2007 & Minor improvements based on feedback\\
    1.1.3   & Nov 23, 2007 & Add OpenGL extension handling\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

During window and pipe initialization the event handling is set up. For
both entities, \textsf{initEvent\-Handler} is called to register the
pipe or window with an event handler. This method may be overwritten to
use a custom event handler, or to not install an event handler at all
and so disable event handling. Likewise, \textsf{exitEventHandler} is
called to de-initialize event handling.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/eventInit.pdf}
  {\caption{\small\label{fEventInit}Event Handling Initialization}}
\end{wrapfigure}
An event handler consists of two parts: the generic base class providing
the interface and generic functions, and the window-system-specific
part providing the actual implementation.

@@

\subsection{OpenGL Extension Handling}

Equalizer uses GLEW\footnote{\link{http://glew.sourceforge.net}} for
OpenGL extension handling, particularly the GLEW MX implementation for
multi-context support.

Each \textsf{eq::Window} has a \textsf{GLEWContext}. This context can be
obtained by using \textsf{glewGetContext} on the window or channel. GLEW
MX uses this function to dispatch the functions to the correct
context. Equalizer (re-)initializes the GLEW context whenever a new
OpenGL context is set on the window.

Extended OpenGL functions called from a window or channel instance can
be called directly. GLEW will call the object's \textsf{glewGetContext}
to obtain the correct context:

{\footnotesize\begin{lstlisting}
void eqPly::Channel::_drawModel( const Model* model )
{
    ...
    glUseProgram( program );
    ...
}
\end{lstlisting}}

Functions called from another place need to have the macro or function
\textsf{glewGetContext} that returns the pointer to the GLEWContext of
the appropriate window.

{\footnotesize\begin{lstlisting}
// state has GLEWContext* from window
#define glewGetContext state.glewGetContext

/*  Set up rendering of the leaf nodes.  */
void VertexBufferLeaf::setupRendering( VertexBufferState& state,
                                       GLuint* data ) const
{
    ...
    glBindBuffer( GL_ARRAY_BUFFER, data[VERTEX_OBJECT] );
    glBufferData( GL_ARRAY_BUFFER, _vertexLength * sizeof( Normal ),
                    &_globalData.normals[_vertexStart], GL_STATIC_DRAW );
    ...
}
\end{lstlisting}}

@@

The main idea for direct send is to parallelize the costly recomposition
for database (sort-last) decomposition. With each additional source
channel, the amount of pixel data to be composited grows linearly. When
using the simple approach of compositing all frames on the destination
channel, this channel quickly becomes the bottleneck in the
system. Direct send distributes this workload evenly across all source
channels, and thereby keeps the compositing work per channel constant.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\small\label{fDirectSend}Parallel Direct Send Compositing}}
\end{wrapfigure}
The following operations have to be possible in order to perform this
algorithm:
\begin{itemize}
\item Selection of color and/or depth frame buffer attachments
\item Restricting the read back area to a part of the rendered area
\item Positioning the pixel data correctly on the receiving channels
\end{itemize}

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/assembly.pdf}
  {\caption{\small\label{fAssembly}Hierarchy of assembly classes}}
\end{wrapfigure}
The frame synchronization through the frame data allows the input frame
to wait for the pixel data to become ready, which is signalled by the
output frame after readback.

Furthermore, the frame data transports the inherited range of the output
frame's compound. The range can be used to compute the assembly order of
multiple input frames, e.g., for sorted-blend compositing in volume
rendering applications.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/volResult.png}
  {\caption{\small\label{fVolResult}Result of \fig{fBlendPerspective}}}
\end{wrapfigure}
Finding the correct assembly order for perspective frustra is more
complex. The perspective distortion invalidates a simple orientation
criteria like the one used for orthographic frustra. For the view and
frustum setup shown in \fig{fBlendPerspective}\footnote{Volume Data Set
  courtesy of: AVS, USA} the correct compositing order is 4-3-1-2 or
1-4-3-2.

In order to compute the assembly order, \textsf{eVolve} uses the angle
between the origin$\rightarrow$slice vector and the near plane, as shown
in \fig{fBlendPerspective}. When the angle becomes greater than
90\textdegree, the compositing order of the remaining frames has to be
changed. The result image of this composition naturally looks the same
as the volume rendering would when rendered on a single
channel. \fig{fVolResult} shows the result of the composition from
\fig{fBlendPerspective}.
